A data-driven model of nucleosynthesis with chemical tagging in a lower-dimensional latent space

### 2019-09-11 article/reviewer-response-v0.txt

RESPONSE SUMMARY:

 1. Done!
 2. Done!
 3. Done!
 4. Need to make a schematic diagram and write response.
 5. Done!
 6. Done!
 7. Done!
 8. Need to justify S/N choice --> text will depend on #12.
 9. Need to add text about N ~ 1000 stars in the toy model (awaiting figure to finish)
10. Done!
11. Done!
12. Waiting for eval-5.py to finish, then consider re-running with S/N > 40.
13. Done!
14. Done!

@@

1.  REVIEWER:

    Abstract:
    The authors mention they use the 1,072 stars GALAH dataset with 18 chemical abundances yet in section 3.2 (where they discuss the dataset) they actually only list 17 elements. In addition only 17 elements are also shown in Fig. 5, 6, 9. That is to say are there 18 or 17 elements that the authors consider?

    RESPONSE:

    There were, and are now, 17 elements. We have dropped V (as per #14) and added Mg. Updated text throughout.

@@

    Can the authors comment on what would happen if the "true" underlying distribution of in latent space is not in fact Gaussian?

    RESPONSE:

    We have added a paragraph that addresses this point in Section 4 (Discussion). That paragraph reads:

        There are similar limitations that arise due to our assumption about the clustering in latent space.
        There is no justified reason why the factor scores should be well-described by multivariate normal
        distributions. If the \emph{true} underlying scores were not distributed as multivariate normals
        then one can imagine similar outcomes when directly fitting data with a mixture of gaussian distributions:
        additional components would be required to describe complex (non-gaussian) shapes in data space. 
        This situation of model mismatch is more extreme when fitting only data rather than the model
        described here because some of the data complexity will be described by the orthogonal latent
        factors. However, qualitatively the picture is the same: when the \emph{true} underlying distribution
        in factor scores are not described by multivariate normals, additional components will likely be
        introduced in order to describe non-gaussian features.

@@

    Section 2:
    The authors state : "... there is a lower dimensional latent space in which the data are clustered, and that clustering is projected into real space by common factor loads" 
    The authors also mention that factor loads can be thought of as the mean yields of a nucleosynthetic event.

    Though we also know that not all Type Ia or Type II nucleosynthetic events can be thought of as the same (i.e. the nucleosynthetic events may not actually be clustered at all). Thus does one expect, given a predicted set yields, that there should be clustering in this latent space?

    RESPONSE:

    Conditioned on our other assumptions explicitly described in the manuscript, we would expect there to be 
    clustering in latent space unless the yields produced by every nucleosynthetic mechanism (including Type Ia supernovae, Type II supernovae, AGB stars, etc) are entirely stochastic.

    Even if Type Ia and Type II supernovae differ in the yields per event, there will still be some mean yield from each type of event and some intrinsic scatter around that mean. The differing means and intrinsic scatter would appear as clustering in latent space. If the intrinsic scatter among each type of event were so large that event types could not be distinguished then we would recover only one (broad) component in latent space.

@@

    If we consider higher S/N cuts then the sample size reduces. As a consequence, fewer latent factors and components are required to explain those data. Of the factors that are found, they are quantitatively similar to those that are found from using the larger sample of stars. In short: the results stay the same, but using a larger sample of stars means we recover more factors.

    Indeed, one of the reasons we started with small numbers of stars (of high quality) and increased to a larger sample was to demonstrate that similar latent factors are found (Figure 9).

    As a general note: We have updated our code to allow for observational errors, but in its current
    implementation this greatly increases the computational cost (by orders of magnitude). We are
    investigating ways that this can be improved, but this is outside the scope of this paper.

    %%% ARC also note this is S/N > 40 in blue. what does that correspond to elsewhere?

@@

9.  REVIEWER:

    Section 3.1 :
    The authors build a toy model in which they sample 100,000 'stars' using 20 clusters in a 5-D factor load space. The toy model seems to work (as per Fig.1). However when they move on to real data in section 3.3 they first only use a sample size of 1,000 (1% the sample size in the toy model) stars. It would be good for the authors to recompute their toy model sampling N=100,000 and N=1,000 'stars' and add this to Figure 1 to better illustrate the effect of lowering the sample size.

    RESPONSE:

    We have updated Figure 1 to show the grid search using N = 100,000 'stars' and N = 1,000 'stars'. We have also updated the text (see PDF changes).

    % ARC comment on changes

@@

13. REVIEWER:

    Fig. 8:
    It would be good to show all 17 elements rather than only the 15 elements shown by the authors.

    RESPONSE:

    This was a bug in our plotting code. Figure 8 has been updated to include all 17 elements (16 as [X/Fe] on y-axis and [Fe/H] as common x-axis).

@@

    However, Buder et al. 2018 states "For V, we see a significant disagreement with both the dwarf abundances and APOGEE DR14. We report this element anyway, because of the useful abundances in dwarfs, but advise not to use of [V/Fe] in giants."

    Are the authors largely using giant stars in the 1,072 stars from GALAH or is this sample dominated by dwarfs? If they there are any giant stars within these 1,072 stars the authors should disregard the V abundances as per Buder et al. 2018.

    RESPONSE:

    The sample included many giant stars. We have removed V from the set of abundances used and updated the text accordingly.
